---
title: "Initial Korean covid 19"
author: "Chi N"
date: "summer2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

As we look at the data, the appropriate variables to use would be either sex that is all lowercase or sex with the uppercase 5 and 1 feel that the appropriate one would be the sex with the uppercase because the variable are just F and M. The second variable that would apply to this would be the result because the only ones you can use are alive or dead, two options. 




## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

As we look at the data, we see that there are more women than men in the alive graph. The dead pile show more man that were dead than female. The alive pile represents approximately 55% people that were female and approximately that were male. When looking at the dead graph, it shows  approximately 37% that were female while approximately 63% that were male .Even though there were more females that caught Covid the ones dying were the female 2-1.

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

When looking at these results, we see that of the dead people, about more than twice are male. At 1.4% of females were dead whereas 3% of males were dead. This showing that it was killing males at double the rate than females. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The females were from a population of a total of 1831 shows that 25 dies whereas in the males population of 1455, 44 died. There wasn't a difference between the sexes of who was getting killed. In the fishers exact tests, the odds of dying if you are a man is 2,3 times more than if you were a woman. 

## Conclusions

As we entered the codes we got score unexpected results. This showed us that sex did not play a role in who was dying. There were more females than male in the graphical results but thst did not matter. Males died at a rate of 2.3 times faster rate than the female. The total population of men were 1856 and omly 25 died. To conclude the results were not what we expected. 
